---
title: "Data Cleaning - Global Temperature"
author: "Katherine Mederos, Claire Nguyen, Meghna Joshi"
date: "2025-11-11"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(here)  # handles paths reliably
source(here("r","00_requirements.R"))
knitr::opts_chunk$set(echo = TRUE)

```

##Introduction 

This R Markdown file performs **data cleaning** for the Global Land-Ocean from 1880-2025. The goals of this file are:

1. Inspect the raw data to comprehend the structure.
2.Clean the data by renaming columns, modifying types, and fix missing values to clean the data.
3. Store the clean data in RDS and CSV files to examine.
4. Provide preliminary visualization in global temperatures trends.


```{r clean-data}
#Load required packages and functions
library(here)
source(here("r", "00_requirements.R"))
source(here("r","01_funct_DataCleaning.R"))
source(here("r","03_funct_Plots.R")) #for plotting


#First 10 lines of raw CSV for structure 
raw_lines<- readLines(here("data", "GlobalTemperatures.csv"), n =10)

#Clean the data using function
clean_temp<-clean_global_temp(here("data", "GlobalTemperatures.csv"))

#Save the cleaned data(CSV/RSD)
save_cleaned_data(clean_temp,"output")

#Inspect cleaned out data
head(clean_temp)
summary(clean_temp)

#Plot cleaned temperature data
plot_global_temp(clean_temp)
```
